DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-01-26 rebuttal.tex

\usepackage{longtable}
\usepackage[LGR,T1]{fontenc}
\usepackage{lscape}
%\usepackage{pdfsync}
\usepackage{multirow}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{lastpage}
\usepackage{afterpage}
\usepackage{lettrine}
\usepackage{color,soul}
\usepackage[dvipsnames]{xcolor}
\usepackage{colortbl}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{titlesec}

%Palatino font
%\usepackage{pxfonts}
%\usepackage{libertine}
\usepackage[scaled=0.88]{beraserif}
\usepackage[scaled=0.85]{berasans}
\usepackage[scaled=0.84]{beramono}
\usepackage{mathpazo}
%\linespread{1.05}
\usepackage[T1,small,euler-digits]{eulervm}

\usepackage[nomessages]{fp}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{bm}

@@

\global\mdfdefinestyle{manuscript}{%
     linecolor=gray!20,linewidth=0.05cm,backgroundcolor=gray!20,%
     leftmargin=-0.5cm,rightmargin=-0.5cm, innerleftmargin=0.4cm,innerrightmargin=0.4cm
}

  \renewcommand{\subsectionautorefname}{Comment}

\newcommand{\nati}[1]{{\color[rgb]{.1,.6,.1}{NP: #1}}}
\newcommand{\TK}[1]{{\color{red}{TK: #1}}}
\newcommand{\todo}[1]{{\color[rgb]{.6,.1,.6}{TODO: #1}}}

\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
\newcommand{\secref}[1]{Section~\ref{sec:#1}}
%\newcommand{\secref}[1]{\S\ref{sec:#1}}
\newcommand{\eqnref}[1]{(\ref{eqn:#1})}

\renewcommand{\b}[1]{{\bm{#1}}}   % bold symbol

% MATH SYMBOLS
\newcommand{\1}{\b{1}}              % all-ones vector
\newcommand{\0}{\b{0}}              % all-zero vector
\newcommand{\g}[1]{\b{#1}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\V}{\mathcal{V}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\B}{\mathcal{B}}
\renewcommand{\L}{\b{L}}
\newcommand{\tL}{\tilde{\L}}
\newcommand{\W}{\b{W}}
\newcommand{\I}{\b{I}}
\newcommand{\D}{\b{D}}
\newcommand{\U}{\b{U}}
\newcommand{\x}{\b{x}}
\newcommand{\X}{\b{X}}
\newcommand{\y}{\b{y}}
\newcommand{\Y}{\b{Y}}
\newcommand{\bu}{\b{u}}
\newcommand{\f}{\b{f}}
\newcommand{\trans}{^\intercal}
\newcommand{\R}{\mathbb{R}}
\newcommand{\bLambda}{\b{\Lambda}}
\newcommand{\blambda}{\b{\lambda}}
\newcommand{\bO}{\mathcal{O}}
\newcommand{\T}{\mathcal{T}}
\DeclareMathOperator*{\esp}{E}
\DeclareMathOperator*{\var}{Var}
\DeclareMathOperator*{\vect}{vec}
\DeclareMathOperator*{\argmin}{arg \, min}
\newcommand{\pkg}[1]{\texttt{#1}}

@@

I will happily recommend this paper for publication. Firstly though, I would like access to the data (privately rather than via emailing the group) so that I can check the notebooks thoroughly. The notebooks are extremely well written but I would like to run them first. As well as this I think there are a few extra applications which will improve the manuscript further, and make the method more appealing for users. I would recommend that these ideas are at least seriously considered by the authors and hopefully implemented: 
\end{mdframed}

\todo{Assigned @Tomek
\begin{itemize}
  \item Thank the reviewer
  \item give him access to data
\end{itemize}}

\subsection{}

\begin{mdframed}[style=comment] 
The first, and most important is an example using masks. Since the authors mention how their method should in principle work well with masks, especially in comparison to HEALPix and Clebsh-Gordan transform spherical harmonic methods I think that it would be extremely enlightening to present this work, probably in the form of a performance vs. time plot. With this being done, the informational content of a masked version of the cosmological application could easily be included. This would, I'm sure, be attractive to the cosmological community as well as the ML community who are always looking for good ways to deal with masked data.
\end{mdframed}
\todo{Assigned: @Michael, @Nati}
Our implementation is actually already working with a mask, we choose to use parts of $1/12$, $1/48$ and $1/192$ of the total sphere. Our main motivation was to build a "large" dataset with a relatively low number of cosmological maps. 
In term of neural network, the availlable implementation of \cite{cohen2018sphericalcnn,kondor2018clebsch} for spherical CNNs and Clebsh-Gordan CNNs are not scalable enough to be trained on the cosmological maps dataset and do not use the HEALPix sampling. Hence a comparison with thee methods requires a different setting, which far beyond the scope of our contribution. Nevertheless, this comparizon is important and we preparing a second contribution where it will be conducted.

\nati{We could compare the price of one convolution. What do you think? We can write the theoretical complexity. This is a sensitive point! }

@@

\subsection{}

\begin{mdframed}[style=comment] 
Secondly, using only SVM as a comparison might skew how well this technique appears to works in a deep learning setting. In principle (and I say this not knowing for sure) 2D CNNs should probably be able to work as well as DeepSphere (although no where near as elegantly) by taking in a projected 2D image of the sky and using *loads* of filters to learn about the distortion at different parts of the projected sky. It would be interesting (and I think a fairly easy test) to check how well DeepSphere performs against a 2D CNN on projections. 
\end{mdframed}
\todo{Assigned: @Nati, @Michael, @Tomek}

\nati{\textbf{Very important point!!!}
\begin{itemize}
  \item We will do this comparizon...
  \item Since we used 1/12 of the sphere, all the pixel grids we use are somehow squared. I am now using this as a projection. @Tomek, do you know if this is ok? Or do you know how to do the projection in a simple and efficient manner? 
  \item I think that traditional 2D CNNs will be as good as our algorithm. We should be very careful in the way we formulate this in our answer and in the paper.
\end{itemize}
}

\begin{mdframed}[style=comment] 
Similarly, I think it would useful to check how well the DeepSphere CNN does against the FCN when the input data is rotated randomly during training - I imagine that the results should match or probably exceed the FCN results and I think this would be a neat little test which doesn't involve needing extra data, just more augmentation.
\end{mdframed}
\todo{Assigned: @Nati, @Michael, @Tomek}

Theoretically, providing that the FCN and the CNN have the same modelization power (i.e, the FCN does not have more or larger layers) and assuming the convolution would be Truly invariant to rotation, the CNN could never exceed the FCN for truly rotational invariant data. However, assuming enought augmentation (as suggested by the reviewer) and training, the CNN should get to the level of the FCN. In practice, this would require more training as the CNN has to learn the invariance. 
The point of defining a FCN instead of a CNN is to avoid the complexity of making this augmention. Hence we did not implement it. Nevertheless, it is an interesting point worth discussion in the paper. We updated the manuscript accordingly.
\todo{Do it}

\nati{
Technically, testing this is complicated as rotation on HEALPix are complicated and expensive. We probably will have to break a lot of code to do it.

@Michael: what do you think?
}

@@

\subsection*{Minor comment}

\begin{mdframed}[style=comment] 
I should say, even though I made the above comments, I thought the paper was really excellent and well written. There were a couple of very minor points which I would like the authors to expand slightly on or consider.
\end{mdframed}
Thank you very much for your comments. We adressed them to the best we could.

@@

\subsection{}
\begin{mdframed}[style=comment] 
While our construction will not exactly respect the setting defined by [33], we observe empirically strong evidence of convergence.

- Could you add a plot comparing the graph Laplacian as a function of number of pixels and the Laplace-Beltrami operator so the reader can explicitly see the convergence?
\end{mdframed}

This is a good question and a fair request. Problematically, we cannot provide such a plot without solving the theoretical problem first. Indeed to experimentally show convergence, we need to actually build a series of graph Laplacian that converges toward the Laplace Beltrami operator. This far more complicated than building one Laplacian and observe some similarities between the two operators.

This is actually a problem we are currently investigating. We plan to make a separate contribution.

@@

\subsection{}

\begin{mdframed}[style=comment] 
We found out that the one proposed above works well for our purpose, and did not investigate other approaches, leaving it to future work.

- Are you really going to search through different weighting schemes in future work? I would be interested if you were, but could you speculate on what you expect to achieve by looking at different ones?
\end{mdframed}

Yes we are investigating the question of graph weights.
As suggested in the previous, the weighting scheme used is very important for our laplacian to approach the true Laplace Beltrami operator. Using a more appropriate weighting scheme, one could hope for a more rotational invariant convolution and a faster convergence rate of the opertor. 
\nati{@michael, how much shall we say here?}

@@

\subsection{}
\begin{mdframed}[style=comment] 
Thanks to the softmax, the output $\bf{y}\in\mathbb{R}^{N_{classes}}$ of a neural network is a probability distribution over the classes, i.e., $y_i$ is the probability that the input sample belongs to class $i$.

- This is a common misconception which is incorrect and needs restating. It is *not* a probability distribution for the class, rather it is a discretised conditional distribution for the class given both the data and the weights and biases of the trained network. One way to see that the value of $y_i$ cannot be considered a probability (which implies P(y|x), i.e. the probability of a class given the data) is that if the network is untrained, would you retrieve back the correct probability? It's wholly dependent on the network.
\end{mdframed}

We agree with the reviewer. We actually misformulate the this sentence. We have reformulate it as:
\begin{mdframed}[style=manuscript] 
Thanks to the softmax, the output $\y \in \R^{N_{classes}}$ of a neural network is a probability distribution over the classes, i.e., $y_i$ is the confidence of the network that the input sample belongs to class $i$. Note that in general, this probability distribution does not correspond to the true posterior distribution $p(y|x)$.
\end{mdframed}

\nati{By the way, this mean that the reviewer is knowledgable with ML, and may be Bayesian. He might be working on another spherical approach.}

@@

\subsection{}
\begin{mdframed}[style=comment] 
Than can be seen as intrinsic data augmentation, as a CNN would need to see many rotated versions of the same data to learn the invariance.

- I think you mean "This" rather than "Than". I think, however, you are making a more important point here than you realise. It is not really "intrinsic data augmentation" but rather building a NN to suit your purposes. This should be done by the entire community as, at the moment, too many people blindly build their networks without considering their data and then struggle to get the best results out. The symmetries of the data should be the first thing to influence architecture of a network, exactly as you have reasoned. You might want to add a statement about how you know what your symmetries are and thus, you know how to build a well reasoned network which is more likely to work (and that other should do the same).
\end{mdframed}
\todo{Assigned: @michael}

\subsection{}
\begin{mdframed}[style=comment] 
As such, the SHT is only performed once on the input (no back and forth SHT between each layer). While this clever trick lowers the complexity of the convolution to $\mathcal{O}(N_{pix})$, the non-linearity is $\mathcal{O}(N^{3/2}_{pix})$.

- I don't understand what the difference is between the convolution and the "non-linearity". I think this sentence either needs explaining further. Is the process $\mathcal{O}(N_{pix})$ or $\mathcal{O}(N^{3/2}_{pix})$?
\end{mdframed}
In the work of \cite{kondor2018clebsch}, the convolution is perfomed in the spectral (harmonic) domain with a simple Adamar product. Instead of recomposing the signal and applying a traditional non-linearity such as "relu" in the pixel domain, \cite{kondor2018clebsch} propose to apply directly a Clebsch–Gordan product in the spectral domain. As a non-linear operation preserving the rotation equivariance, it is considered as the non-linearity of the neural network.

We have modified the manuscript as:
\begin{mdframed}[style=manuscript] 
While this clever trick lowers the complexity of the convolution (Adamar product in the spectral domain) to $\bO(N_{pix})$, the non-linearity (Clebsch–Gordan product) is $\bO(N_{pix}^{3/2})$.
\end{mdframed}
\nati{Should we be more clear in the paper?}

\subsection{}
\begin{mdframed}[style=comment] 
When mentioning that you are 10 to 20 times faster than the SHT it should be noted that libsharp has amazing distributed mpi functionality which helps, although of course your method is genuinely lower order computationally (and can be distributed simply on a GPU).
\end{mdframed}
When comparing the speed, we used a single core CPU implemention of the graph convolution. We believe that this comparizon is fair.
\nati{Is this sufficent?}

@@

\subsection{}
\begin{mdframed}[style=comment] 
In Appendix A. (my personal favourite part) could you measure the percent difference that different eigenvectors differ from the spherical harmonics due to incomplete discretisation of the sphere, perhaps plotting side by side plots like Figure 3, but with SH too (or their differences). By the way, Figure A.12 is placed in the wrong section.
\end{mdframed}
\todo{Answer + compute the ratio of outside block energy (see Fig. A.12)}

\subsection{}
\begin{mdframed}[style=comment] 
In Appendix C. is there a connection between $t$ and $\sigma$ that you can make, or do you just fit them to be similar? If you just fit them to be similar I'm not sure where the relative differences come from. If they are related, however, then the relative differences make sense, but I can't see what the connection is.
\end{mdframed}

In the current version of the manuscript and the code, we simply fitted a few $t$ to match the different $\sigma$. In theory, there should be a connection between the two values providing the graph Laplacian has converged toward the Laplace Beltrami operator. Again, this is left for a futur contribution.

\subsection{}
\begin{mdframed}[style=comment] 
A question that I would like to know personally, are you thinking about a future upgrade to non-radial kernels?
\end{mdframed}
We are thinking about it. However, as the graph framework we are using only works with radial filters, it is not a simple question. One simple trick to make network sensitive to direction is to add a per channel fully connected layers. We are currently investigating other potential graph techniques that would lead to non-radial kernels.

\subsection{Details}
\begin{mdframed}[style=comment] 
Finally, I noticed a few points in the text where there were typos or things which were difficult to follow and I have put a note of them below and tried to correct them as much as I can so that you don't have to put too much effort in - I can't wait to see this paper published. Sorry that the list is so extensive (I'm a bit of a stickler for details and trying to write papers concisely):
\end{mdframed}
\todo{Assigned @tomek. 
This reviewer really spend the time. We should say something.}

@@

\begin{mdframed}[style=comment] 
Convolutional Neural Networks (CNNs) have been proposed as an alternative analysis tool in cosmology thanks to their ability to automatically design relevant statistics to maximise the precision of the parameter estimation [9–16].

- There are many of us in the community who are not overly keen on "precision" of parameter estimation being the target for machine learning. Rather, we would prefer robustness of parameter estimation. It might be worth changing this sentence a bit so as not to upset some of the referenced authors ;)
\end{mdframed}
\todo{Assigned: @tomek, @michael}

\begin{mdframed}[style=comment] 
When you have multiple = in equations could you split them on to separate lines? This would make the logical steps for following the equations a lot clearer.
\end{mdframed}
\todo{Assigned: @tomek}

\begin{mdframed}[style=comment] 
You change a lot between "node", "vertex" and "pixel". Could you stick to one of these? I think that "vertex" or "pixel" makes the most sense ("pixel" due to the HEALPix pixelisation of the maps or vertex because of the use of the $v$ notation). Node is quite confusing in several places because there is an inherent connection between node and neuron in ML literature, and since there is swapping between the three different terms it's a little hard to keep track.
\end{mdframed}
\todo{Assigned: @all}
\nati{What do you think? This avoid repetition. But we could be more consistent...}

\begin{mdframed}[style=comment] 
You change between math font and text font for the names of your layers when inside and outside of equations. It would look a lot neater sticking with one choice. (Sorry, that seems quite petty and I don't mean it to be).
\end{mdframed}
Thank you for the suggestion. We have made the paper consistent.
\nati{@all, when you read the paper, please check that I did not forget some of them. Every layer has be changed to math symbole as we use them like function. For example SM becomes $SM$.}

\begin{mdframed}[style=comment] 
In Appendix C. you talk about Kroneker a lot, but earlier you talk about Kroneker $\delta$ (which is it's correct name). It's probably worth sticking with the full name.
\end{mdframed}
We have modified the manuscript according to the reviewer suggestion.

\begin{mdframed}[style=comment] 
Sky maps are rotation invariant: rotating maps on the sphere doesn’t change their interpretation, as only the statistics of the maps are relevant.
- I think that you mean that rotating maps on the sphere doesn't change their interpretation *when* only the statistics of the maps are relevant. Say you wanted to compare the NGC against the SGC, then rotating the map would obviously change the results.
\end{mdframed}
This sentence has been reformulated as:
\begin{mdframed}[style=manuscript] 
We consider the sky maps rotationial invariant as in our case rotating maps on the sphere doesn't change their interpretation since only the statistics of the maps are relevant.
\end{mdframed}

@@

\begin{mdframed}[style=comment] 
The flexibility of modelling the data domain with a graph allows to easily model data that spans only a part of the sphere, or data that is not uniformly sampled.

- I think you're missing a "one" after allows. "The flexibility of modelling the data domain with a graph allows one to easily model data that spans only a part of the sphere, or data that is not uniformly sampled."
\end{mdframed}
\todo{assigned: @tomek}
\nati{I am not sure of the english here.}

@@

\begin{mdframed}[style=comment] 
A CNN is composed of the following main building blocks [31]: (i) a convolution, (ii) a non-linearity, (iii) a down-sampling operation, (iv) a pooling operation, and (v), optionally, normalization.

- I think that this is generalising a CNN a bit too much. In my mind (and in my personal uses of CNNs) steps (iii) and (iv) are not always necessary either. Perhaps you want to say that you wish to tackle common tasks which often arise in CNNs: i) convolution, (ii) non-linearity, (iii) down-sampling operation, (iv) pooling operation, and (v), optionally, normalisation.
\end{mdframed}
\todo{assigned: @michael}

\begin{mdframed}[style=comment] 
Likewise, down-sampling can be achieved by taking one pixel out of n.

### 2019-01-26 scnn.tex
going further in the rebuttal

So far these algorithms have mostly been demonstrated on Euclidean domains, such as images.
The main challenge in designing a CNN on the sphere is to define a convolution operation that is suitable for this domain, while taking care of the necessary irregular sampling.
Moreover, the designed convolution and resulting neural network should possess the following three key characteristics.
First, the convolution should be equivariant to rotation, meaning that a rotation of the input implies the same rotation of the output.
Sky maps are rotation invariant: rotating maps on the sphere doesn't change their interpretation, as only the statistics of the maps are relevant.
Depending on the task, we want the CNN to be either equivariant or invariant to rotation.
Second, to be able to train the network in reasonable time, the convolution has to be computationally efficient.
Third, a CNN should work well on parts of the sphere, as many cosmological observations cover only a part of the sky.
For ground-based observations this can be due to limited visibility of the sky from a particular telescope location, and for space-based instruments due to masking of the galactic plane area (see \figref{example_maps} for example maps).

@@

We give a practical demonstration of DeepSphere on cosmological model discrimination using maps of projected mass distribution on the sky \citep{chang2017curvedsky}.
This kind of maps can be created using the gravitational lensing technique \citep[see][for review]{BartelmannSchneider2001weak}.
Our maps are similar to the ones used by \citep{schmelze2017cosmologicalmodel}.
In a simplified scenario, we classify partial sky convergence maps into two cosmological models.
These models were designed to have very similar angular power spectrum.
We compare the performance of DeepSphere to a baseline algorithm: an SVM classifier that takes pixel histograms or power spectrum densities (PSDs) of these maps as input.
The comparison is made as a function of the additive noise level and the area of the sphere used in the analysis.
Results show that our model is always better at discriminating the maps, especially in the presence of noise.
DeepSphere is implemented with \pkg{TensorFlow} \citep{abadi2016tensorflow} and is intended to be easy to use out-of-the-box for cosmological applications.
Code and examples are available online.{\interfootnotelinepenalty=10000 \footnote{\url{https://github.com/SwissDataScienceCenter/DeepSphere/}}}

@@

On regular Euclidean domains, such as 1-dimensional time series or 2-dimensional images, a convolution can be efficiently implemented by sliding a localized convolution kernel (for example a patch of $5 \times 5$ pixels) in the signal domain.
Likewise, down-sampling can be achieved by taking one pixel out of $n$.
Because of the irregular sampling, there is no straightforward way to define a convolution on the sphere directly in the pixel domain; convolutions are most often performed using spherical harmonics.
In our method we also use the spectral domain to define the convolution.
The implementation, however, does not need a direct access to the spectrum, which is computationally more efficient (see \secref{efficient_convolution}).
The gist of our method is to define the convolution operation on a sphere using a graph, and the down-sampling operation using a hierarchical pixelisation of the sphere.

@@

Contrary to classical signal processing on Euclidean domains, the kernel $h$ has no single representation in the vertex domain and cannot be translated on the graph. It can however be \textit{localized} on any vertex $v_i$ by the convolution with a Kronecker delta\footnote{A Kronecker delta is the signal $\b \delta^i \in \R^{N_{pix}}$ that is zero everywhere except on vertex $v_i$ where it takes the value one.} $\b \delta^i \in \R^{N_{pix}}$. The localization operator $\T_i$ reads $\T_i h = h(\L) \b \delta^i = (h(\L))_i$, the $i$th column of $h(\L)$.
This localization of the kernel $h$ can be useful to visualize kernels, as explained in~\ref{sec:filter_visualization}.
If the graph is not regular, i.e., all nodes do not have the same number of neighbors, and all distances are not equal, the effect of the kernel will slightly differ from one node to another. While there is no perfect sampling of the sphere, these differences are negligible as the structure of the whole graph is very regular. However, when considering only parts of the sphere, one can observe important border effects (see \figref{border_effects} and~\ref{sec:border_effects}).

Finally, the graph convolution can be interpreted in the vertex domain as a scalar product with localizations $\T_i h$ of the kernel $h$. Indeed, the result of the convolution of the signal $\f$ with the kernel $h$ is
\begin{equation} \label{eqn:graph_convolution_spatial}
	(h(\L) \f)_i = \langle \T_i h(\L), \f \rangle = \langle h(\L) \b \delta^i, \f \rangle.
\end{equation}
To make the parallel with the classical 1D convolution, let $f, g\in \mathbb{Z} \rightarrow \mathbb{R}$ be two 1D discrete signals. Their convolution can be written in the same form as \eqnref{graph_convolution_spatial}:
\begin{equation*}
	(f \ast g) [i] = \sum_{j=-\infty}^\infty f[j] g[i-j] = \langle T_i g,  f \rangle,
\end{equation*}
where $T_i g[j] = g[i-j]$ is, up to a flip, a translation operator.
Similarly as \eqnref{graph_convolution_spatial}, the convolution of the signal $f$ by a kernel $g$ is the scalar product of $f$ with translated versions $T_i g$ of the kernel $g$.
Additionally, it turns out that the localization operator $\T_i h$ is a generalization of the translation operator on graphs. In the particular case where the Laplacian matrix $\L$ is circulant, $\T_i h$ is a translated version of $\T_j h$ for all $i, j$ and both convolutions are equivalent. We refer the reader to \citep[Sec 2.2]{perraudin2017stationary} for a detailed discussion of the connection between translation $T_i$ and localization $\T_i$.

@@

Fortunately, both of these computational issues are overcome by defining the convolution kernel $h$ as a polynomial $h_\theta(\lambda) = \sum_{k=0}^K \theta_k \lambda^k$ of degree $K$ parametrised by $K+1$ coefficients $\theta$.
The filtering operation \eqnref{graph_convolution_fourier} becomes
\begin{equation} \label{eqn:graph_convolution_monomial}
	h_\theta(\L) \f =  \U \left(\sum_{k=0}^K \theta_k \bLambda^k \right) \U\trans \f = \sum_{k=0}^K \theta_k \L^k \f,
\end{equation}
where $\L^k$ captures $k$-neighborhoods.\footnote{The 1-neighborhood of a node is the set of nodes that are directly connected to it. The $k$-neighborhood of a node is the set of nodes that are connected to it through paths of length $k$.}
The entry $(\L^k)_{ij}$ is the sum of all weighted paths of length $k$ between nodes $v_i$ and $v_j$, where the weight of a path is the multiplication of all the edge weights on the path.
Hence, it is non-zero if and only if nodes $v_i$ and $v_j$ are connected by at least one path of length $k$.
Filtering with a polynomial convolution kernel can thus be interpreted in the pixel (vertex) domain as a weighted linear combination of neighboring pixel values.
In the classical setting, convolutions are also weighted local sums.
The weights are however given by the filter coefficients only, and there is one coefficient per pixel in the patch.
In the graph setting, the weights are determined by the filter coefficients $\theta$ and the Laplacian $\L$, and there is one coefficient per neighborhood, not node.
That defines radial filters, values of which depend only on the distance to the center, and not on the direction.
% It is required as there is no way to order and match nodes from the same neighborhood.
% The number of nodes in a neighborhood most often even vary from node to node.
While it may seem odd to restrict filters to be 1D while the sphere is 2D, radial filters are direction-less and result in rotation equivariant convolutions.
We note that restricting the graph convolutional kernel to a polynomial is similar to restricting the classical Euclidean convolution to a fixed-size patch.
Similarly, each line of the matrix $\sum_{k=0}^K \theta_k \L^k$ defines an irregular patch of radius $K$.
Hence, filters designed as polynomials of the Laplacian have local support in the vertex domain.

>>>>

So far these algorithms have mostly been demonstrated on Euclidean domains, such as images.
The main challenge in designing a CNN on the sphere is to define a convolution operation that is suitable for this domain, while taking care of the necessary irregular sampling.
Moreover, the designed convolution and resulting neural network should possess the following three key characteristics.
First, the convolution should be equivariant to rotation, meaning that a rotation of the input implies the same rotation of the output.
\nati{We consider the sky maps rotationial invariant as in our case rotating maps on the sphere doesn't change their interpretation since only the statistics of the maps are relevant.}
Depending on the task, we want the CNN to be either equivariant or invariant to rotation.
Second, to be able to train the network in reasonable time, the convolution has to be computationally efficient.
Third, a CNN should work well on parts of the sphere, as many cosmological observations cover only a part of the sky.
For ground-based observations this can be due to limited visibility of the sky from a particular telescope location, and for space-based instruments due to masking of the galactic plane area (see \figref{example_maps} for example maps).

@@

We give a practical demonstration of DeepSphere on cosmological model discrimination using maps of projected mass distribution on the sky \citep{chang2017curvedsky}.
These kind of maps can be created using the gravitational lensing technique \citep[see][for review]{BartelmannSchneider2001weak}.
Our maps are similar to the ones used by \citep{schmelze2017cosmologicalmodel}.
In a simplified scenario, we classify partial sky convergence maps into two cosmological models.
These models were designed to have very similar angular power spectrum.
We compare the performance of DeepSphere to a baseline algorithm: an SVM classifier that takes pixel histograms or power spectrum densities (PSDs) of these maps as input.
The comparison is made as a function of the additive noise level and the area of the sphere used in the analysis.
Results show that our model is always better at discriminating the maps, especially in the presence of noise.
DeepSphere is implemented with \pkg{TensorFlow} \citep{abadi2016tensorflow} and is intended to be easy to use out-of-the-box for cosmological applications.
Code and examples are available online.{\interfootnotelinepenalty=10000 \footnote{\url{https://github.com/SwissDataScienceCenter/DeepSphere/}}}

@@

On regular Euclidean domains, such as 1-dimensional time series or 2-dimensional images, a convolution can be efficiently implemented by sliding a localized convolution kernel (for example a patch of $5 \times 5$ pixels) in the signal domain.
\nati{Likewise, down-sampling can be achieved by keeping one pixel every $n$ pixels with $n=2$ being very common.}
Because of the irregular sampling, there is no straightforward way to define a convolution on the sphere directly in the pixel domain; convolutions are most often performed using spherical harmonics.
In our method we also use the spectral domain to define the convolution.
The implementation, however, does not need a direct access to the spectrum, which is computationally more efficient (see \secref{efficient_convolution}).
The gist of our method is to define the convolution operation on a sphere using a graph, and the down-sampling operation using a hierarchical pixelisation of the sphere.

@@

Contrary to classical signal processing on Euclidean domains, the kernel $h$ has no single representation in the vertex domain and cannot be translated on the graph. It can however be \textit{localized} on any vertex $v_i$ by the convolution with a Kronecker delta\footnote{A Kronecker delta is the signal $\b \delta^i \in \R^{N_{pix}}$ that is zero everywhere except on vertex $v_i$ where it takes the value one.} $\b \delta^i \in \R^{N_{pix}}$. The localization operator $\T_i$ reads $\T_i h = h(\L) \b \delta^i = (h(\L))_i$, the $i$th column of $h(\L)$.
This localization of the kernel $h$ can be useful to visualize kernels, as shown in an example of heat diffusion presented in~\ref{sec:filter_visualization}.
If the graph is not regular, i.e., all nodes do not have the same number of neighbors, and all distances are not equal, the effect of the kernel will slightly differ from one node to another. While there is no perfect sampling of the sphere, these differences are negligible as the structure of the whole graph is very regular. However, when considering only parts of the sphere, one can observe important border effects (see ~\ref{sec:border_effects}).

Finally, the graph convolution can be interpreted in the vertex domain as a scalar product with localizations $\T_i h$ of the kernel $h$. Indeed, the result of the convolution of the signal $\f$ with the kernel $h$ is
\begin{equation} \label{eqn:graph_convolution_spatial}
	(h(\L) \f)_i = \langle \T_i h(\L), \f \rangle = \langle h(\L) \b \delta^i, \f \rangle.
\end{equation}
To make the parallel with the classical 1D convolution, let $f, g\in \mathbb{Z} \rightarrow \mathbb{R}$ be two 1D discrete signals. Their convolution can be written in the same form as \eqnref{graph_convolution_spatial}:
\begin{equation*}
	(f \ast g) [i] = \sum_{j=-\infty}^\infty f[j] g[i-j] = \langle T_i g,  f \rangle,
\end{equation*}
where $T_i g[j] = g[i-j]$ is, up to a flip ($g[i-j]$ instead of $g[i+j]$), a translation operator.
Similarly as \eqnref{graph_convolution_spatial}, the convolution of the signal $f$ by a kernel $g$ is the scalar product of $f$ with translated versions $T_i g$ of the kernel $g$.
Additionally, it turns out that the localization operator $\T_i h$ is a generalization of the translation operator on graphs. In the particular case where the Laplacian matrix $\L$ is circulant, $\T_i h$ is a translated version of $\T_j h$ for all $i, j$ and both convolutions are equivalent. We refer the reader to \citep[Sec 2.2]{perraudin2017stationary} for a detailed discussion of the connection between translation $T_i$ and localization $\T_i$.

@@

Fortunately, both of these computational issues are overcome by defining the convolution kernel $h$ as a polynomial $h_\theta(\lambda) = \sum_{k=0}^K \theta_k \lambda^k$ of degree $K$ parametrised by $K+1$ coefficients $\theta$.
The filtering operation \eqnref{graph_convolution_fourier} becomes
\begin{equation} \label{eqn:graph_convolution_monomial}
	h_\theta(\L) \f =  \U \left(\sum_{k=0}^K \theta_k \bLambda^k \right) \U\trans \f = \sum_{k=0}^K \theta_k \L^k \f,
\end{equation}
where $\L^k$ captures $k$-neighborhoods.\footnote{The 1-neighborhood of a node is the set of nodes that are directly connected to it. The $k$-neighborhood of a node is the set of nodes that are connected to it through paths of length $k$.}
The entry $(\L^k)_{ij}$ is the sum of all weighted paths of length $k$ between nodes $v_i$ and $v_j$, where the weight of a path is the multiplication of all the edge weights on the path.
Hence, it is non-zero if and only if nodes $v_i$ and $v_j$ are connected by at least one path of length $k$.
Filtering with a polynomial convolution kernel can thus be interpreted in the pixel (vertex) domain as a weighted linear combination of neighboring pixel values.
In the classical setting, convolutions are also weighted local sums.
The weights are however given by the filter coefficients only, and there is one coefficient per pixel in the patch.
In the graph setting, the weights are determined by the filter coefficients $\theta$ and the Laplacian $\L$, and there is one coefficient per neighborhood, not node.
That defines radial filters, values of which depend only on the distance to the center, and not on the direction.
% It is required as there is no way to order and match nodes from the same neighborhood.
% The number of nodes in a neighborhood most often even vary from node to node.
While it may seem odd to restrict filters to be 1D while the sphere is 2D, radial filters are direction-less and result in rotation equivariant convolutions.
We note that restricting the graph convolutional kernel to a polynomial is similar to restricting the classical Euclidean convolution to a fixed-size patch.
Similarly, each column of the matrix $\sum_{k=0}^K \theta_k \L^k$ defines an irregular patch of radius $K$.
Hence, filters designed as polynomials of the Laplacian have local support in the vertex domain.

Following \citep{defferrard2016convolutional}, we define our filters as Chebyshev polynomials.
The filtering operation \eqnref{graph_convolution_fourier} becomes
\begin{equation} \label{eqn:graph_convolution_cheby}
	h_\theta\left(\tL\right) \f = \U \left(\sum_{k=0}^K \theta_k T_k\left(\tilde{\bLambda}\right) \right) \U\trans \f = \sum_{k=0}^K \theta_k T_k\left(\tL\right) \f,
\end{equation}
where 
\begin{equation*}
\tL = \frac{2}{\lambda_{max}} \L - \b{I} = -\frac{2}{\lambda_{max}} \D^{-1/2} \W \D^{-1/2}
\end{equation*}
is the rescaled Laplacian with eigenvalues $\tilde{\b \Lambda}$ in $[-1, 1]$.
$T_k(\cdot)$ is the Chebyshev polynomial of degree $k$ defined by the recursive relation $T_k(\tL) = 2\tL T_{k-1}(\tL) - T_{k-2}(\tL)$, $T_1(\tL) = \tL$, $T_0(\tL) = \b{I}$.
While definitions \eqnref{graph_convolution_monomial} and \eqnref{graph_convolution_cheby} both allow the representation of the same filters, we found in our experiments that optimizing $\theta$ in \eqnref{graph_convolution_cheby} is slightly more stable than $\theta$ in \eqnref{graph_convolution_monomial}.
We believe that this is due to (i) their almost orthogonality in the spectral and spatial domains, and (ii) their uniformity.\footnote{The amplitude of the Chebyshev polynomials $T_k(x)$ is mostly constant over the domain $[-1, 1]$, independently of the order $k$. On the contrary, the amplitude of the monomials $x^k$ is very different for $|x|\approx 0$ and $|x| \approx 1$.}
Finally, note that while the graph convolution \eqnref{graph_convolution_fourier} is motivated in the spectral domain, definitions \eqnref{graph_convolution_monomial} and \eqnref{graph_convolution_cheby} are implementations in the vertex domain.

Exploiting the recursive formulation of Chebyshev polynomials, evaluating \eqnref{graph_convolution_cheby} requires $\bO(K)$ multiplications of the vector $\f$ with the sparse matrix $\tL$.
The cost of one such multiplication is $\bO(|\E| + |\V|)$.
By construction of our graph, $|\E| < 8 N_{pix}$ and the overall computational cost of the convolution reduces to $\bO(N_{pix})$ operations and as such is much more efficient than filtering with spherical harmonics, even though HEALPix was designed as an iso-latitude sampling that has a fast spherical transform.
This is especially true for smooth kernels which require a low polynomial degree $K$.
\figref{filtering_speed} compares the speed of low-pass filtering for Gaussian smoothing using the spherical harmonics and the graph-based method presented here.
A naive implementation of our method is ten to twenty times faster for $N_{side} = 2048$, with $K=20$ and $K=5$, respectively, than using the spherical harmonic transform at $\ell_{max} = 3 N_{side}$ implemented by the highly optimized \pkg{libsharp} \citep{reinecke2013libsharp} library used by HEALPix.
